---
title: '070223'
author: "Joaquin Peraza"
date: "2023-02-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Frequentist apporach

$y = 0$

$[y|p] = Bern(p)$

$\vec{y} = (y_1,y_2,y_3,...)$

$[\vec{y}|p] = \Pi Bern(p)$

$\hat{p} = \frac{1}{n} \sum y$

$\hat{p} = \frac{\sum y:+\frac{1}{2}}{n+1}$

Bayesian approadch add more information to the model

$[y|p] = Bern(p)$

$[\vec{y}|p] = \Pi Bern(p)$

$[p] = Beta(\alpha=1, \beta=1)$

Every probability is possible with equaly likely to occur

Fitting the model

$[p|y] = \frac{[\underline{y}|p] \cdot[p]}{\int  [\underline{y}|p][p] dp}$

$[p|y] = Beta(\tilde{\alpha}, \tilde{\beta})$

$\tilde{\alpha} = \alpha + \sum y:$

$\tilde{\beta} = \beta + n -\hat{\sum} y:$


Posterior distribution 

$\alpha = 1$

$\beta = 2$


```{r}
y <- 0
n <- length(y)

alpha <- 1
beta <- 1

K <- 5000
alpha.tilde <- alpha+sum(y)
beta.tilde <- beta+n-sum(y)

p.y <- rbeta(K, alpha.tilde, beta.tilde)

hist(p.y, freq = FALSE, xlab = "p|y", ylab = "[y|p]")

samples <- matrix(,K,1)
samples[1,] <- 0.01

for (k in 2:K){
  p.old <- samples[k-1,]
  p.try <- runif(1,0,1)
  mh <- (prod(dbinom(y,1,p.try))*dbeta(p.try,alpha,beta)) /
        (prod(dbinom(y,1,p.old))*dbeta(p.old,alpha,beta))
  keep <- ifelse(mh>1,1,rbinom(1,1,mh))
  samples[k,] <- ifelse(keep==1, p.try, p.old)
}

hist(samples, freq = FALSE, xlab = "p|y", ylab = "[y|p]")



samples <- matrix(,K,1)
samples[1,] <- 1/10^7

for (k in 2:K){
  p.old <- samples[k-1,]
  p.try <- runif(1,1/10^9, 1/10^6)
  mh <- (prod(dbinom(y,1,p.try))*dunif(p.try,1/10^9, 1/10^6)) /
        (prod(dbinom(y,1,p.old))*dunif(p.old,1/10^9, 1/10^6))
  keep <- ifelse(mh>1,1,rbinom(1,1,mh))
  samples[k,] <- ifelse(keep==1, p.try, p.old)
}

hist(samples, freq = FALSE, xlab = "p|y", ylab = "[y|p]")
samples
```