---
title: "Assignment 1"
author: "Joaquin Peraza"
date: "2023-01-17"
output: html_document
---

1. Take 500 draws from a normal distribution that has an expected value of 2 and a variance of 9. Make a histogram of these 500 draws.

```{r}
draws<-rnorm(500, mean = 2, sd = sqrt(9))
hist(draws)
```

2. Calculate the sample mean from the 500 draws from problem 1.

```{r}
mean(draws)
```

3. Take 500 draws from Bernoulli distribution that has an expected value of 0.5. Make a histogram of these 500 draws.

```{r}
draws<-rbinom(500, 1,.5)
hist(draws)
```

4. Calculate the sample mean from the 500 draws from problem 3.

```{r}
mean(draws)
```

5. Take 500 draws from Bernoulli distribution that has a variance of 0.25.

$$
\sigma^2 = p(1-p) \\
0.25 = p(1-p) \\
p = 0.5
$$

```{r}
draws<-rbinom(500, 1,.5)
```

6. Take 500 draws from a normal distribution that has an expected value of 0 and a variance of 1. Raise these 500 draws to the second power and make a histogram.


```{r}
draws<-rnorm(500, mean = 0, sd = sqrt(1))
draws_power <- draws^2
hist(draws_power)
```

7. What well-known and named probability density function is shown in the histogram you made in question 6.


|       **Exponential distribution**
